---
title: "Assignment 7"
author: "Veronika Palotai"
date: "11/2/2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Exercise 1**

All done although setting up an existing directory as a project was a bit different from creating a completely new one. The necessary step were 'File' -> 'New Project' -> 'Existing Directory'.

**Exercise 2**

All done.

**Exercise 3**

Import Libraries

```{r, packages_install, eval=TRUE, results='hide', message=FALSE, warning=FALSE}
library(ggplot2)
library(tidyverse)
library(scales)
```

- Import from csv

Setting the Path

```{r, path, eval=TRUE, results='hide', warning=FALSE, message=FALSE}
dir <-  "D:/Egyetem/CEU/Coding_1/R-Coding/"
```

Location Folders

```{r, loc, eval=TRUE, results='hide', warning=FALSE, message=FALSE}
data_in <- paste0(dir,"da_data_repo/hotels-europe/clean/")
```

Loading Datasets 

```{r, dataset, eval=TRUE, results='hide', warning=FALSE, message=FALSE}
hotels_prices_csv <- read_csv(paste0(data_in,"hotels-europe_price.csv"))
hotels_features_csv <- read_csv(paste0(data_in,"hotels-europe_features.csv"))
```


- Import from .dta

Import 'haven' library and look up `read_dta`

```{r, hav, eval=TRUE, warning=FALSE, message=FALSE, results='hide'}
library(haven)
?read_dta
```

Reading .dta files

```{r, dta, eval=TRUE, message=FALSE, warning=FALSE, results='hide'}
hotels_prices_dta <- read_dta(paste0(data_in,"hotels-europe_price.dta"), encoding = NULL)
hotels_features_dta <- read_dta(paste0(data_in,"hotels-europe_features.dta"), encoding = NULL)
```

- Comparing contents

First, let's look at the contents of the tables.

```{r, comp, eval=TRUE, message=FALSE, warning=FALSE}
View(hotels_features_dta)
View(hotels_features_csv)
View(hotels_prices_dta)
View(hotels_prices_csv)
```

We can see that the order of the fields and the records are different. Let's check the length of each table to see whether they contain equal number of observations.

```{r, length, eval=TRUE, message=FALSE, warning=FALSE}
hotels_features_csv %>%
  summarise(
    nr_of_records = n()
  )

hotels_features_dta %>%
  summarise(
    nr_of_records = n()
  )

hotels_prices_csv %>%
  summarise(
    nr_of_records = n()
  )

hotels_prices_dta %>%
  summarise(
    nr_of_records = n()
  )
```

As we can see, they contain the same number of records. Let's go into the details a bit and check the number of records for each city in the features table and each year in the prices table.

```{r, length2, eval=TRUE, message=FALSE, warning=FALSE}
hotels_features_csv %>%
  group_by(city) %>%
  summarise(
    nr_of_records = n()
  )

hotels_features_dta %>%
  group_by(city) %>%
  summarise(
    nr_of_records = n()
  )

hotels_prices_csv %>%
  group_by(year) %>%
  summarise(
    nr_of_records = n()
  )

hotels_prices_dta %>%
  group_by(year) %>%
  summarise(
    nr_of_records = n()
  )
```

Most of the time they are the same in both tables, however, we can see a difference in case of 'Amsterdam' which might be due to a reading error. Since the overall number of records are the same, there has to be at least one other city with a difference.

**Exercise 4**

Picking the first 200 lines of the prices table

```{r, subset, eval=TRUE, message=FALSE, warning=FALSE}
(hotels_prices_csv_small <- hotels_prices_csv %>%
  slice(1:200))
```

Writing to file

```{r, wr, eval=TRUE, message=FALSE, warning=FALSE}
write.csv(hotels_prices_csv_small, file = "hotels-europe-small.csv", row.names = FALSE)
```

- Lines changed: I deleted the field names

- It lead to the problem that the contents of the first row were interpreted as column names

- Fixing:

```{r, re, eval=TRUE, message=FALSE, warning=FALSE}
hotels_small <- read.csv("hotels-europe-small.csv", header = FALSE)
# names columns as V1, V2, V3, ...
```

**Exercise 5**

12.6 Case Study

Load the dataset:

```{r, case, eval=TRUE, warning=FALSE, message=FALSE}
who <- tidyr::who
```

Gather together all the columns from `new_sp_m014` to `newrel_f65` because they are values not variables:

```{r, gather, eval=TRUE, message=FALSE, warning=FALSE}
who1 <- who %>% 
  gather(new_sp_m014:newrel_f65, key = "key", value = "cases", na.rm = TRUE)
```

Some hint on the structure of the values in the new key column by counting them:

```{r, cnt, eval=TRUE, message=FALSE, warning=FALSE}
who1 %>% 
  count(key)
```

Minor fix to the format of the column names:

```{r, minfix, eval=TRUE, message=FALSE, warning=FALSE}
who2 <- who1 %>% 
  mutate(key = stringr::str_replace(key, "newrel", "new_rel"))
```

Separate the values in each code with two passes of `separate()`. The first pass will split the codes at each underscore.

```{r, sep2, eval=TRUE, message=FALSE, warning=FALSE}
who3 <- who2 %>% 
  separate(key, c("new", "type", "sexage"), sep = "_")
```

Drop the 'new' column because it’s constant in this dataset. Let’s also drop 'iso2' and 'iso3' since they’re redundant.

```{r, drop, eval=TRUE, message=FALSE, warning=FALSE}
who3 %>% 
  count(new)

who4 <- who3 %>% 
  select(-new, -iso2, -iso3)
```

Separate 'sexage' into 'sex' and 'age' by splitting after the first character:

```{r, sep, eval=TRUE, message=FALSE, warning=FALSE}
who5 <- who4 %>% 
  separate(sexage, c("sex", "age"), sep = 1)
```

The 'who' dataset is now tidy, however, we could have done all these steps at the same time, like this:

```{r, together, eval=TRUE, message=FALSE, warning=FALSE}
who %>%
  gather(key, value, new_sp_m014:newrel_f65, na.rm = TRUE) %>% 
  mutate(key = stringr::str_replace(key, "newrel", "new_rel")) %>%
  separate(key, c("new", "var", "sexage")) %>% 
  select(-new, -iso2, -iso3) %>% 
  separate(sexage, c("sex", "age"), sep = 1)
```

**Exercise 6**

The dataset:

```{r, data_exercise_t, eval=TRUE, message=FALSE, warning=FALSE}
(df <- tibble(name = c("A123", "B456"), age = c(30, 60), answer1 = c(0, 1), answer2 = c(1,1), answer3 = c(1,0), answer4 = c(0,0)))
```

1. Cleaning by hand - description only 

What needs to be done is to gather the columns 'answers1' to 'answers4' and create an 'answer_nr' (key) and an 'answer_value' (value) column for the answers and their corresponding values pertaining to a specific age group.

2. Cleaning by coding it with `gather()`

```{r, code, eval=TRUE, message=FALSE, warning=FALSE}
tidy_df <- df %>%
  gather(answer1:answer4, key="answer_nr", value="answer_value")
tidy_df
```